Constant Partying: \\ Introducing a Generic Toolkit for Recursive Partitioning in R

### 2014-02-10 pkg/partykit/inst/NUS-2014/constparty_TH_AZ.Rnw
start constant party talk
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/partykit@454 255f4249-2263-4417-b1ea-8498779e27a7

\documentclass[11pt,compress,t]{beamer}
\usetheme{Z}
\usepackage{amsfonts,amstext,amsmath}
%% need no \usepackage{Sweave}
\definecolor{InputColor}{rgb}{0,0,0.3}
\definecolor{OutputColor}{rgb}{0.2,0.2,0.2}

%% commands
\newcommand{\ui}{\underline{i}}
\newcommand{\oi}{\overline{\imath}}
\newcommand{\darrow}{\stackrel{\mbox{\tiny \textnormal{d}}}{\longrightarrow}}
\newcommand{\parrow}{\stackrel{\mbox{\tiny \textnormal{p}}}{\longrightarrow}}
\newcommand{\dotequals}{\stackrel{\cdot}{=}}
\newcommand{\efp}{\mathrm{\it efp}}
\newcommand{\given}{\, | \,}
\newcommand{\ltime}{\lambda_\mathrm{time}}
\newcommand{\lcomp}{\lambda_\mathrm{comp}}
\newcommand{\argmin}{\operatorname{argmin}\displaylimits}

@@

<<packages>>=
library("partykit")
library("rpart")
library("RWeka")
library("evtree")
@

@@

<<pmml, eval=FALSE>>=
pm <- pmmlTreeModel("ttnc.xml")
@

@@

\title{partykit: A Toolkit for Recursive Partytioning}
\author{Achim Zeileis, Torsten Hothorn}
\URL{http://eeecon.uibk.ac.at/~zeileis/}
\lecture{partykit: A Toolkit for Recursive Partytioning}

\subsection{Overview}

@@

\begin{itemize}
  \item Status quo: R software for tree models
  \item New package: partykit
  \begin{itemize}
    \item Unified infrastructure for recursive partytioning
    \item Classes and methods
    \item Interfaces to \code{rpart}, \code{J48}, \dots
    \item Illustrations
  \end{itemize}
  \item Future: Next steps
\end{itemize}

@@

\subsection{Status quo: R software for tree models}

@@

\subsection{New package: partykit}

@@

<<rpart-print, echo=TRUE, results=verbatim>>=
rp
@

@@

<<rpart-print, echo=TRUE, results=verbatim>>=
as.party(rp)
@

@@

\textbf{Prediction:} Compare rpart's C code and partykit's R code
for (artificially) large data set.

@@

\textbf{PMML:} Predictive Model Markup Language. XML-based format
exported by various software packages including SAS, SPSS, R/pmml.
Here, reimport CART tree.

@@

\begin{frame}
\frametitle{partykit: PMML}

@@

\subsection{Future: Next steps}

\begin{frame}
\frametitle{Next steps}

\textbf{To do:} The current code is already fairly well-tested and mostly
stable. However, there are some important items on the task list.
\begin{itemize}
  \item Extend/smooth package vignette.
  \item Add \code{xtrafo}/\code{ytrafo} to \fct{ctree} reimplementation.
  \item Switch \fct{mob} to new \class{party} class.
  \item Create new subclass \class{modelparty} that facilitates
    handling of formulas, terms, model frames, etc.\ for model-based trees.
\end{itemize}

@@

\begin{frame}[fragile]
\frametitle{Next steps}

@@

\textbf{Illustration:} Logistic regression (or logit model), assessing differences in
the effect of ``preferential treatment'' for women or children.

<<mob-show, echo=TRUE, eval=FALSE>>=
<<mob>>
@

Odds ratio of survival given treatment differs across subsets (slope), as does
the survival probability of male adults (intercept).

<<mob-coef, echo=TRUE, results=verbatim>>=
coef(mb)
@

@@

\begin{frame}
\frametitle{Next steps}

\vspace*{-0.3cm}

\setkeys{Gin}{width=1.1\textwidth}
\hspace*{-0.4cm}%
<<Titanic-mob-visualization, fig=TRUE, height=7, width=12>>=
plot(mb, tp_args = list(margins = c(1.5, 2, 1.5, 2.5)))
@

\subsection{Computational details}

@@

\textbf{Data:} The contingency table \code{Titanic} is transformed
to the data frame \code{ttnc} using the following code. 
%
<<data1, echo=TRUE, eval=FALSE>>=
<<data>>
@
%
The last transformation also adds the treatment variable for the model-based recursive
partitioning analysis.

@@

\subsection{References}

>>>>

\documentclass[11pt,compress,t]{beamer}

\usetheme{uzhneu-en-informal}
\usepackage[T1]{fontenc}
\usepackage[utf8]{inputenc}
\usepackage{amsfonts,amstext,amsmath,amssymb,amsthm}

%% need no 
\usepackage{Sweave}
\definecolor{InputColor}{rgb}{0,0,0.3}
\definecolor{OutputColor}{rgb}{0.2,0.2,0.2}

%% commands
\let\code\texttt
\let\class\texttt
\let\fct\texttt
\let\doi\texttt
\newcommand{\dquote}[1]{``#1''}
\newcommand{\ui}{\underline{i}}
\newcommand{\oi}{\overline{\imath}}
\newcommand{\darrow}{\stackrel{\mbox{\tiny \textnormal{d}}}{\longrightarrow}}
\newcommand{\parrow}{\stackrel{\mbox{\tiny \textnormal{p}}}{\longrightarrow}}
\newcommand{\dotequals}{\stackrel{\cdot}{=}}
\newcommand{\efp}{\mathrm{\it efp}}
\newcommand{\given}{\, | \,}
\newcommand{\ltime}{\lambda_\mathrm{time}}
\newcommand{\lcomp}{\lambda_\mathrm{comp}}
\newcommand{\argmin}{\operatorname{argmin}\displaylimits}

@@

<<packages>>=
library("partykit")
library("Formula")
library("rpart")
library("RWeka")
library("evtree")
@

@@

<<pmml, eval=FALSE>>=
pm <- pmmlTreeModel(file.path(system.file("pmml", 
    package = "partykit"), "ttnc.pmml"))
@

@@

\title[Constant Partying]{Constant Partying: \\ 
      Introducing a Generic Toolkit for Recursive Partitioning in R}
\author{Torsten Hothorn (Universit\"at Z\"urich) \\
        Achim Zeileis (Universit\"at Innsbruck)}
%\URL{http://user.math.uzh.ch/hothorn/}
%\lecture{partykit: A Toolkit for Recursive Partytioning}

\maketitle

@@

\begin{itemize}
  \item Status quo: R software for tree models
  \item New package: partykit
  \begin{itemize}
    \item Unified infrastructure for recursive partytioning
    \item Classes and methods
    \item Interfaces to \code{rpart}, \code{J48}, \dots
    \item Illustrations
  \end{itemize}
  \item Here and now: Intro to partykit for trees with constant fits in each node.
  \item Future: Next steps
\end{itemize}

@@

@@

@@

\begin{frame}[fragile]
\frametitle{partykit: Interface to \texttt{rpart}}
\small{
<<rpart-print, echo=TRUE, results=verbatim>>=
rp
@
}

@@

\small{
<<rpart-print, echo=TRUE, results=verbatim>>=
as.party(rp)
@
}

@@

\textbf{Prediction:} Compare rpart's C code and partykit's R code

@@

\textbf{PMML:} Predictive Model Markup Language. XML-based format
exported by various software packages including SAS, SPSS, R/pmml.
Here, import QUEST tree.

@@

\begin{frame}
\frametitle{partykit: QUEST via PMML}

@@

\begin{frame}
\frametitle{Your own tree}

@@

\begin{frame}
\frametitle{What's next?}

@@

Talk by Achim; but I'm entitled to a coffee first...

@@

@@

\textbf{Data:} The contingency table \code{Titanic} is transformed
as explained in the \code{constparty} package vignette.

@@
